---
title: "Assignment 2 - Meta-analysis of pitch in schizophrenia"
author: "Maria Olsen"
date: "16/8/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
pacman::p_load(brms,gridExtra,msm)#tidybayes,


```

# Assignment 2: meta-analysis

## Questions to be answered

1. 
Simulate data to setup the analysis and gain insight on the structure of the problem. 

Simulate one dataset of 100 studies 

(n of participants should follow a normal distribution with mean of 20, sd of 10, but no fewer than 10 participants), 

with a mean effect size of 0.4, average deviation by study of .4 and measurement error of .8. 

The data you get should have one row per study, with an effect size mean and standard error. 

Build a proper bayesian model to analyze the simulated data. 

Then simulate publication bias (only some of the studies you simulate are likely to be published, which?), 

the effect of publication bias on your estimates (re-run the model on published studies, assess the difference), and discuss what this implies for your model. remember to use at least one plot to visualize your results. 


BONUS question: do a power/precision analysis: w this kind of sample sizes (participants) how many studies would you need to acquire good precision (e.g. .1 sd in the pop level estimate)

2. What is the current evidence for distinctive vocal patterns in schizophrenia? 
Use the data from Parola et al (2020) - https://www.dropbox.com/s/0l9ur0gaabr80a8/Matrix_MetaAnalysis_Diagnosis_updated290719.xlsx?dl=0 - focusing on pitch variability (PITCH_F0SD).  Describe the data available (studies, participants). Using the model from question 1 analyze the data, visualize and report the findings: population level effect size; how well studies reflect it; influential studies, publication bias. 
BONUS question: assess the effect of task on the estimates (model comparison with baseline model)


# Question 1

#trying again
```{r}
#setting parameters
study_ES <- 0.4
study_sd <- 0.4
error <- 0.8
n_studies <- 100

#making dataframe

d <- tibble(
  Study_ID = seq(n_studies),
  SampleSize = round(rtnorm(n_studies, 20, 10, lower= 10)),
  StudyEffect = NA,
  ES_mean = NA,
  ES_SE = NA,
  Published = NA,
  PublishedPos = NA
)

for (i in seq(n_studies)){
  d$StudyEffect[i] <- rnorm(n_studies, study_ES, study_sd)
  sampling <- rnorm(d$SampleSize[i], d$StudyEffect[i], error)
  d$ES_mean[i] <- mean(sampling)
  d$ES_SE[i] <- sd(sampling)/sqrt(d$SampleSize[i])
  d$Published[i] <- ifelse(
    abs(d$ES_mean[i]) - (2*d$ES_SE[i])>0,
        rbinom(1, 1, 0.9), rbinom(1, 1, 0.1))
  d$PublishedPos[i] <- ifelse(
    abs(d$ES_mean[i]) - (2*d$ES_SE[i])>0 & d$ES_mean[i] > 0,
        rbinom(1, 1, 0.9), rbinom(1, 1, 0.1))
}

view(d)


index                          <- d$Study_ID+1
d[index:(index+2),]            <- NA
d$Study_ID[index:(index+2)]       <- c(index:(index+2))
d$SampleSize[index:(index+2)]  <- c(25,30,27)
d$StudyEffect[index:(index+2)]  <- study_ES
d$ES_mean[index:(index+2)]    <- c(2.5,3,2.7)
d$ES_SE[index:(index+2)] <- 1
d$Published[index:(index+2)]   <- 1
d$PublishedPos[index:(index+2)]<- 1



hist_all <- ggplot(d) +
  aes(x = ES_mean) +
  geom_histogram(bins = 30L, fill = "#4682B4") +
  labs(title = "Effect mean of the studies, all studies") +
  geom_vline(xintercept = 0, color="black") +
  theme_minimal() +
  theme(plot.title = element_text(size = 18L, face = "bold"))


#making subset of only publicated based on the positive publication bias
d_pub <- d %>% 
  subset(d$PublishedPos == "1")


hist_pub <- ggplot(d_pub) +
  aes(x = ES_mean) +
  geom_histogram(bins = 30L, fill = "#4682B4") +
  labs(title = "Effect mean of the studies, published studies") +
  geom_vline(xintercept = 0, color="black") +
  theme_minimal() +
  theme(plot.title = element_text(size = 18L, face = "bold"))

hist_all
hist_pub

ggsave("hist_of_all_studies.png",hist_all)
ggsave("hist_of_only_published.png",hist_pub)
```
```{r}
d$PublishedPos <- as.factor(d$PublishedPos)

sim_data_plot <- ggplot(d) +
  aes(
    x =  reorder(Study_ID, -StudyEffect),
    y = StudyEffect,
    colour = PublishedPos,
    group = PublishedPos
  ) +
  labs(x="Study")+
  geom_jitter(size = 1.5) +
  theme_minimal()

ggsave("sim_data_plot.png", sim_data_plot)

sim_data_plot
```



```{r}
#making model
m <- ES_mean| se(ES_SE) ~ 1 + (1|Study_ID)

get_prior(m, data = d, family= gaussian)

#Making priors
prior_meta <- c(
  prior(normal(0, 0.3), class= Intercept),
  prior(normal(0, 0.2), class= sd)
)

mp_all <- brm(
    m, 
    data = d,
    family = gaussian,
    prior = prior_meta,  
    sample_prior = "only", 
    iter = 5000,
    warmup = 1000,
    backend = "cmdstanr",
    threads = threading(2),
    cores = 2,
    chains = 2,
    control = list(adapt_delta = 0.99, max_treedepth = 20))


mp_pub <- brm(
    m, 
    data = d_pub,
    family = gaussian,
    prior = prior_meta,  
    sample_prior = "only", 
    iter = 5000,
    warmup = 1000,
    backend = "cmdstanr",
    threads = threading(2),
    cores = 2,
    chains = 2,
    control = list(adapt_delta = 0.99, max_treedepth = 20))


pri_all <- pp_check(mp_all, ndraws=100)+labs("Priors all")
pri_pub <- pp_check(mp_pub, ndraws=100)+labs("Priors pub")

ggsave("prior_check_all.png", pri_all)
ggsave("prior_check_pub.png", pri_pub)

pri_all
pri_pub
```
```{r}
#making posterior
mpp_all <- brm(
    m, 
    data = d,
    family = gaussian,
    prior = prior_meta,  
    sample_prior = T, 
    iter = 5000,
    warmup = 1000,
    backend = "cmdstanr",
    threads = threading(2),
    cores = 2,
    chains = 2,
    control = list(adapt_delta = 0.99, max_treedepth = 20))


mpp_pub <- brm(
    m, 
    data = d_pub,
    family = gaussian,
    prior = prior_meta,  
    sample_prior = T, 
    iter = 5000,
    warmup = 1000,
    backend = "cmdstanr",
    threads = threading(2),
    cores = 2,
    chains = 2,
    control = list(adapt_delta = 0.99, max_treedepth = 20))

post_all <- pp_check(mpp_all, ndraws=100)+labs("Posterior all")
post_pub <- pp_check(mpp_pub, ndraws=100)+labs("Posterior published")

ggsave("posterior_check_all.png", post_all)
ggsave("posterior_check_pub.png", post_pub)

post_all
post_pub

pp_checks <- grid.arrange(pri_all, pri_pub, post_all, post_pub)
ggsave("pp_checks_sim.png", pp_checks)
```


```{r}
samp_all <- as_draws_df(mpp_all)
samp_pub <- as_draws_df(mpp_pub)

variables(mpp_all)

updateplot_intercept_all <- ggplot(samp_all) +
  geom_density(aes(prior_Intercept), fill="steelblue", color="black",alpha=0.6) +
  geom_density(aes(b_Intercept), fill="#FC4E07", color="black",alpha=0.6) + 
  labs(title = "Update plot intercept (all studies)")+
  geom_vline(xintercept = 0.4)+
  theme_classic()

updateplot_sigma_all <- ggplot(samp_all) +
  geom_density(aes(prior_sd_Study_ID), fill="steelblue", color="black",alpha=0.6) +
  geom_density(aes(sd_Study_ID__Intercept), fill="#FC4E07", color="black",alpha=0.6) + 
  labs(title = "Update plot Sigma (all studies)")+
  geom_vline(xintercept = 0.4)+
  theme_classic()


updateplot_intercept_pub <- ggplot(samp_pub) +
  geom_density(aes(prior_Intercept), fill="steelblue", color="black",alpha=0.6) +
  geom_density(aes(b_Intercept), fill="#FC4E07", color="black",alpha=0.6) + 
  labs(title = "Update plot intercept (Pub studies)")+
  geom_vline(xintercept = 0.4)+
  theme_classic()

updateplot_sigma_pub <- ggplot(samp_pub) +
  geom_density(aes(prior_sd_Study_ID), fill="steelblue", color="black",alpha=0.6) +
  geom_density(aes(sd_Study_ID__Intercept), fill="#FC4E07", color="black",alpha=0.6) + 
  labs(title = "Update plot Sigma (Pub studies)")+
  geom_vline(xintercept = 0.4)+
  theme_classic()



#updateplot_intercept_all
#updateplot_sigma_all
#updateplot_intercept_pub
#updateplot_sigma_pub

all_update_plots <- grid.arrange(updateplot_intercept_all, updateplot_sigma_all, updateplot_intercept_pub, updateplot_sigma_pub)

ggsave("all_updateplots_sim.png", all_update_plots)

all_update_plots
```

```{r}
summary(mpp_all)
summary(mpp_pub)
```


```{r}
hypothesis(mpp_pub, "Intercept>0.4")
hypothesis(mpp_all, "Intercept<0.4")
```


## Question 2

#read in data 
```{r}
library(readxl)
data <- read_excel("Matrix_MetaAnalysis_Diagnosis_updated290719.xlsx")
view(data)
```

#escalc (calculate effectsize) 
```{r}
pacman::p_load("metafor")

pitch_varience<-escalc("SMD",
                  n1i=SAMPLE_SIZE_HC, n2i=SAMPLE_SIZE_SZ,
                  m1i=PITCH_F0SD_HC_M,
                  m2i=PITCH_F0SD_SZ_M,
                  sd1i=PITCH_F0SD_HC_SD,
                  sd2i=PITCH_F0SD_SZ_SD,
                  data = data)

view(pitch_varience)

res <- rma(yi, vi, data=pitch_varience) # Perform meta-analysis
metafor::funnel(res)
res # print results
```




#baysian framework
```{r}
real_f <- bf(yi | se(vi) ~ 1 + (1 | Article))

r_priors <- c(
  prior(normal(0,0.36),class=Intercept),
  prior(normal(0,0.8),class=sd))

```

#priors
```{r}
real_prior <- brm(
    real_f, 
    data = pitch_varience,
    family = gaussian,
    prior = r_priors,  
    sample_prior = "only", 
    iter = 5000,
    warmup = 1000,
    backend = "cmdstanr",
    threads = threading(2),
    cores = 2,
    chains = 2,
    control = list(adapt_delta = 0.99, max_treedepth = 20))


pp_prior_real <- pp_check(real_prior, ndraws=100)+labs("Prior check")

ggsave("pp_prior_real.png", pp_prior_real)

pp_prior_real
```

#posterior
```{r}
real_posterior <- brm(
    real_f, 
    data = pitch_varience,
    family = gaussian,
    prior = r_priors,  
    sample_prior = T, 
    iter = 5000,
    warmup = 1000,
    backend = "cmdstanr",
    threads = threading(2),
    cores = 2,
    chains = 2,
    control = list(adapt_delta = 0.99, max_treedepth = 20))

pp_posterior_real <- pp_check(real_posterior, ndraws=100)+labs("Posterior check")

ggsave("pp_posterior_real.png", pp_posterior_real)

pp_posterior_real

```

#updateplots
```{r}
samp_real <- as_draws_df(real_posterior)

variables(real_posterior)

updateplot_intercept <- ggplot(samp_real) +
  geom_density(aes(prior_Intercept), fill="steelblue", color="black",alpha=0.6) +
  geom_density(aes(b_Intercept), fill="#FC4E07", color="black",alpha=0.6) + 
  labs(title = "Update plot intercept")+
  theme_classic()

updateplot_sigma <- ggplot(samp_real) +
  geom_density(aes(prior_sd_Article), fill="steelblue", color="black",alpha=0.6) +
  geom_density(aes(sd_Article__Intercept), fill="#FC4E07", color="black",alpha=0.6) + 
  labs(title = "Update plot Sigma")+
  theme_classic()

updateplots_real <- grid.arrange(updateplot_intercept, updateplot_sigma)
ggsave("updateplots_real.png", updateplots_real)

updateplot_intercept
updateplot_sigma

```
```{r}
summary(real_posterior)
```



#forest plot
https://www.dsquintana.blog/how-to-perform-a-bayesian-meta-analysis-in-r/

```{r}
pacman::p_load(weightr)
weightfunct(pitch_varience$yi, pitch_varience$vi)
```


```{r}
posterior_summary(real_posterior, pars = c("^b_", "^sd_"), probs = c(0.025, 0.975))

post.samples <- posterior_samples(real_posterior, c("^b_", "^sd_"))
names(post.samples)

names(post.samples) <- c("smd", "tau")
names(post.samples)
```

```{r}
smd.ecdf <- ecdf(post.samples$smd)
smd.ecdf(0.3)

#We see that with 0%, the probability of our pooled effect being smaller than 0.30 is very, very low. Assuming the cut-off is valid, this would mean that the overall effect of the intervention we find in this meta-analysis is very likely to be meaningful.
```

#forrest
```{r}
study_draws <- spread_draws(real_posterior, r_Article[Article,], b_Intercept) %>% 
  mutate(b_Intercept = r_Article + b_Intercept)

pooled_eff_draws <- spread_draws(real_posterior, b_Intercept) %>% 
  mutate(Article = "Pooled Effect")

forest_data <- bind_rows(study_draws, 
                         pooled_eff_draws) %>% 
   ungroup() %>%
   mutate(Article = reorder(Article, b_Intercept))

forest_data_summary <- group_by(forest_data, Article) %>% 
  mean_qi(b_Intercept)
```


```{r}
ggplot(aes(b_Intercept, Article 
          ), 
       data = forest_data) +
  
  # Add vertical lines for pooled effect and CI
  geom_vline(xintercept = fixef(real_posterior)[1, 1], 
             color = "grey", size = 1) +
  geom_vline(xintercept = fixef(real_posterior)[1, 3:4], 
             color = "grey", linetype = 2) +
  geom_vline(xintercept = 0, color = "black", 
             size = 1) +
  
  # Adding densities
  geom_density_ridges(fill = "blue", 
                      rel_min_height = 0.01, 
                      col = NA, scale = 1,
                      alpha = 0.8) +
  geom_pointintervalh(data = forest_data_summary, 
                      size = 1) +
  
  # Add text and labels
  geom_text(data = mutate_if(forest_data_summary, 
                             is.numeric, round, 2),
    aes(label = glue("{b_Intercept} [{.lower}, {.upper}]"), 
        x = Inf), hjust = "inward") +
  labs(x = "Standardized Mean Difference", # summary measure
       y = element_blank()) +
  theme_minimal()
```

